---
title: "Walkthough Analysis - Cumulative"
author: "Ben Hicks"
date: "31 October 2017"
output: html_document
---

# Methodology

Walkthroughs conducted from early 2016 to `r Sys.Date()`. The following data was collected from each class:

### Context of Lesson

Question | Description / Method
--------|-----------------------------
period | The period of the lesson visited (1 - 6)
room | Room visited
year level | The year level visited. An answer of "3.5" indicates a mixed class of year 3 and year 4 students
number of students | how many students in the class
lesson time | The time of the lesson, designated as beginning (B), middle (M) or end (E). Beginning was deemed to be anytime in the first 15 minutes (allowing for movement to class) and end was anytime during the last 10 minutes

### Student Activity - What the students were doing

These categories are *not* exclusive. Students could be doing a number of these at once. For instance class discussion with heavy scaffolding from the teacher might be classed as *direct instruction* as well as *discussion*.

Question | Description / Method
--------|-----------------------------
direct instruction | Students were recieving direct instruction from the teacher, or the teacher was modelling work, or similar.
discussing | Students were talking *about what they were learning* or working on.
reading / viewing | Students were reading a text, viewing media, or similar. This also includes listening to a recording.
groupwork | Students were working on a task in groups of two or more. 
independent work | Students were working on a task by themselves. 

### Learning Intentions and Success Criteria

Question | Description / Method
--------|-----------------------------
LI visible | The learning intentions are visible, either on the board, on a document, or online. If online students needed to be able to access / have accessed it.
SC visible | The success criteria are visible, either on the board, on a document, or online. If online students needed to be able to access / have accessed it.

### Student understanding of learning process

This is both the most subjective and important measure of the walkthrough. As such each question was asked to at least 4 students, 2 each for the staff performing the walkthrough. After exiting the class, a short discussion would rate the responses on a scale of 0 - 2, 2 being a *good understanding*, 1 being *some idea*, and 0 being *struggling to understand*. The exact questions would change depending on the student, year level, content, but the goal was to get students to explain what they are learning, and what success in this process would look like. Being a small school it is helpful knowing the students and this was taken into account when evaluating the student responses.

Question | Description / Method
--------|-----------------------------
learning understanding | Questions such as *what are you learning today?* were asked, and students prompted to explain what they were learning about. Scored on a scale of 0 - 2.
success understanding | Questions such as *how do you know when you understand it?* or *what do you hope to get out of the lesson?* were asked. The aim was to prompt students into explaining how successful understanding would be achieved. Scored on a scale of 0 - 2. 

# Results - Today
```{r echo=FALSE} 
# setting up library's and parameters
suppressPackageStartupMessages(library(googleVis))
suppressPackageStartupMessages(library(plyr))
suppressPackageStartupMessages(library(readr))
op <- options(gvis.plot.tag='chart')

# Column names must be in the following sequence
col_names <- c('period','room','year.level','number.of.students','lesson.time','direct.instruction','discussing','reading..viewing','groupwork','independent.work','LI.visible','SC.visible','learning.understanding','success.understanding','date')

new_file <- "data/Walkthrough.2017.10.31.csv"
new_date <- "2017-10-31"
wtd_today <- data.frame(suppressMessages(read_csv(new_file)), date = as.Date(new_date))
names(wtd_today) <- col_names

wtd_previous <- read.csv("data/All.Walkthroughs.csv")
names(wtd_previous) <- col_names
# merging data
wtd01 <- rbind(wtd_today, wtd_previous)
wtd01 <- unique(wtd01)

#####################
#### DO THIS!!!! ####
#####################
# must add export with unique function for All.Walkthroughs
#write.csv(wtd01, file = "data/All.Walkthroughs.csv")
```

### Context

Summary data | Total
-------|----------
Classes visited: | `r length(wtd_today[,1])`
 - - beginning of lesson | `r sum(wtd_today$lesson.time == "B")`
 - - middle of lesson | `r sum(wtd_today$lesson.time == "M")`
 - - end of lesson | `r sum(wtd_today$lesson.time == "E")`
Students observed: | `r sum(wtd_today$number.of.students)`
 - - Primary | `r sum(wtd_today$number.of.students) - sum(subset(wtd_today, wtd_today$year.level %in% c(7,8,9,10,11,12))$number.of.students)`
 - - Secondary | `r sum(subset(wtd_today, wtd_today$year.level %in% c(7,8,9,10,11,12))$number.of.students)`

### Student activity

##### Overall activity breakdown
``` {r echo = FALSE, results = "asis", comment = NA}
# Overall activity breakdown
suppressPackageStartupMessages(library(rCharts))
rd1 <- data.frame(Activity = names(wtd_today[6:10]), Rate = sapply(wtd_today[6:10], mean))
rd1$Activity <- gsub("reading..viewing","reading or viewing",rd1$Activity)
rd1$Activity <- gsub("\\."," ",rd1$Activity)
rd1 <- rd1[order(rd1$Activity),]
p2t <- nPlot(Rate ~ Activity,
            data = rd1,
            type = 'multiBarChart')
p2t$chart(color = c('brown','blue','green','#444444','#aa2288'))
#p2t$templates$script <- "http://timelyportfolio.github.io/rCharts_nvd3_templates/chartWithTitle.html"
p2t$set(title = "Activity breakdown")
p2t$yAxis(tickFormat = "#!d3.format('%')!#")
p2t$print('iframe11',include_assets = TRUE)
```

\n

##### Activity breakdown by time of lesson
``` {r echo = FALSE, results = "asis", comment = NA}
# Activity breakdown on time of lesson
suppressPackageStartupMessages(library(rCharts))

bdat <- subset(wtd_today, wtd_today$lesson.time == "B", select = 6:10)
bmean <- data.frame(lesson.time = "Beginning", lapply(bdat, mean))
mdat <- subset(wtd_today, wtd_today$lesson.time == "M", select = 6:10)
mmean <- data.frame(lesson.time = "Middle", lapply(mdat, mean))
edat <- subset(wtd_today, wtd_today$lesson.time == "E", select = 6:10)
emean <- data.frame(lesson.time = "End", lapply(edat, mean))

bdat <- data.frame(lesson.time = "Beginning",
                   Activity = names(bmean[2:6]),
                   Rate = as.numeric(bmean[1,2:6])
)
mdat <- data.frame(lesson.time = "Middle",
                   Activity = names(mmean[2:6]),
                   Rate = as.numeric(mmean[1,2:6])
)
edat <- data.frame(lesson.time = "End",
                   Activity = names(emean[2:6]),
                   Rate = as.numeric(emean[1,2:6])
)

temp <- merge(bdat,mdat, all = TRUE)
lt.split.act <- merge(temp,edat, all = TRUE)
lt.split.act$Activity <- gsub("reading..viewing","reading or viewing",lt.split.act$Activity)
lt.split.act$Activity <- gsub("\\."," ",lt.split.act$Activity)
lt.split.act.plot <- nPlot(Rate ~ Activity, group = 'lesson.time',
                       type = 'multiBarChart',
                       data = lt.split.act)
#lt.split.act.plot$templates$script <- "http://timelyportfolio.github.io/rCharts_nvd3_templates/chartWithTitle.html"
lt.split.act.plot$set(title = "Activity breakdown based on time of lesson")
lt.split.act.plot$yAxis(tickFormat = "#!d3.format('%')!#")

lt.split.act.plot$print('iframe21',include_assets = TRUE)
```

\n

# Results - Cumulative

```{r echo=FALSE} 

timeSeries <- ddply(wtd01, ~date, summarise,
                    LI = mean(LI.visible),
                    SC = mean(SC.visible),
                    Learning.Understanding = mean(learning.understanding),
                    Success.Understanding = mean(success.understanding))
timeSeries$date <- as.Date(timeSeries$date, "%Y-%m-%d")

```

### Context

Summary data | Total
-------|----------
Classes visited: | `r length(wtd01[,1])`
 - - beginning of lesson | `r sum(wtd01$lesson.time == "B")`
 - - middle of lesson | `r sum(wtd01$lesson.time == "M")`
 - - end of lesson | `r sum(wtd01$lesson.time == "E")`
Students observed: | `r sum(wtd01$number.of.students)`
 - - Primary | `r sum(wtd01$number.of.students) - sum(subset(wtd01, wtd01$year.level %in% c(7,8,9,10,11,12))$number.of.students)`
 - - Secondary | `r sum(subset(wtd01, wtd01$year.level %in% c(7,8,9,10,11,12))$number.of.students)`

### Student activity

##### Overall activity breakdown
``` {r echo = FALSE, results = "asis", comment = NA}
# Overall activity breakdown
suppressPackageStartupMessages(library(rCharts))
rd1 <- data.frame(Activity = names(wtd01[6:10]), Rate = sapply(wtd01[6:10], mean))
rd1$Activity <- gsub("reading..viewing","reading or viewing",rd1$Activity)
rd1$Activity <- gsub("\\."," ",rd1$Activity)
rd1 <- rd1[order(rd1$Activity),]
p2 <- nPlot(Rate ~ Activity,
            data = rd1,
            type = 'multiBarChart')
p2$chart(color = c('brown','blue','green','#444444','#aa2288'))
#p2$templates$script <- "http://timelyportfolio.github.io/rCharts_nvd3_templates/chartWithTitle.html"
p2$set(title = "Activity breakdown")
p2$yAxis(tickFormat = "#!d3.format('%')!#")
p2$print('iframe1',include_assets = TRUE)
```

\n

##### Activity breakdown by time of lesson
``` {r echo = FALSE, results = "asis", comment = NA}
# Activity breakdown on time of lesson
suppressPackageStartupMessages(library(rCharts))

bdat <- subset(wtd01, wtd01$lesson.time == "B", select = 6:10)
bmean <- data.frame(lesson.time = "Beginning", lapply(bdat, mean))
mdat <- subset(wtd01, wtd01$lesson.time == "M", select = 6:10)
mmean <- data.frame(lesson.time = "Middle", lapply(mdat, mean))
edat <- subset(wtd01, wtd01$lesson.time == "E", select = 6:10)
emean <- data.frame(lesson.time = "End", lapply(edat, mean))

bdat <- data.frame(lesson.time = "Beginning",
                   Activity = names(bmean[2:6]),
                   Rate = as.numeric(bmean[1,2:6])
)
mdat <- data.frame(lesson.time = "Middle",
                   Activity = names(mmean[2:6]),
                   Rate = as.numeric(mmean[1,2:6])
)
edat <- data.frame(lesson.time = "End",
                   Activity = names(emean[2:6]),
                   Rate = as.numeric(emean[1,2:6])
)

temp <- merge(bdat,mdat, all = TRUE)
lt.split.act <- merge(temp,edat, all = TRUE)
lt.split.act$Activity <- gsub("reading..viewing","reading or viewing",lt.split.act$Activity)
lt.split.act$Activity <- gsub("\\."," ",lt.split.act$Activity)
lt.split.act.plot <- nPlot(Rate ~ Activity, group = 'lesson.time',
                       type = 'multiBarChart',
                       data = lt.split.act)
#lt.split.act.plot$templates$script <- "http://timelyportfolio.github.io/rCharts_nvd3_templates/chartWithTitle.html"
lt.split.act.plot$set(title = "Activity breakdown based on time of lesson")
lt.split.act.plot$yAxis(tickFormat = "#!d3.format('%')!#")

lt.split.act.plot$print('iframe2',include_assets = TRUE)
```

### Student understanding of learning process

The following charts address the questions asked to students, gauging their understanding of *what they are learning* and *what successful understanding looks like*. The scores have been converted to a percentage for ease of comparison.


``` {r echo = F, results = "asis", warning = F}

# transforming data for charting
t.LI <- data.frame(date = timeSeries$date,
                 Measure = "LI Visible",
                 value = timeSeries$LI)
t.SC <- data.frame(date = timeSeries$date,
                 Measure = "SC Visible",
                 value = timeSeries$SC)
t.LU <- data.frame(date = timeSeries$date,
                 Measure = "Learning Understood",
                 value = timeSeries$Learning.Understanding/2)
t.SU <- data.frame(date = timeSeries$date,
                 Measure = "Success Understood",
                 value = timeSeries$Success.Understanding/2)
t1 <- rbind(t.LI, t.SC)
t2 <- rbind(t.LU, t.SU)
t <- rbind(t1, t2)

#p <- nPlot(value ~ date, data = t, type = "lineChart", group = "measure")
#p$templates$script <- "http://timelyportfolio.github.io/rCharts_nvd3_templates/chartWithTitle.html"
#p$set(title = "Timeline of understanding of learning process")
#p$xAxis( tickFormat="#! function(d) {return d3.time.format('%b %d')(new Date(d*1000*3600*24));} !#" )
#p$yAxis(tickFormat = "#!d3.format('%')!#")
#p$chart(forceY = c(0,1))
#p$print()

# attempting with ggplot instead
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(scales))
suppressPackageStartupMessages(library(plotly))
suppressPackageStartupMessages(library(ggthemes))
ggtime <- ggplot(t, aes(y = value, x = date, color = Measure)) + geom_line()
ggtime <- ggtime + scale_y_continuous(name = "",labels = percent, limits = c(0,1)) + scale_x_date(name = "", labels = date_format("%b-%y"))
ggtime <- ggtime + theme_tufte()
ggtime
```

\n


\n

#### Learning understanding based on lesson conditions
``` {r echo = FALSE, results = "asis", comment = NA}
# Learning Understanding based on categories
# Building data frame
suppressPackageStartupMessages(library(dplyr))
lu.split.livs <- summarise(group_by(wtd01, LI.visible),
                            mean(learning.understanding)+
                             mean(success.understanding))
lu.split.scvs <- summarise(group_by(wtd01, SC.visible),
                            mean(learning.understanding)+
                             mean(success.understanding))
lu.split.drct <- summarise(group_by(wtd01,direct.instruction),
                             mean(learning.understanding)+
                             mean(success.understanding))
lu.split.disc <- summarise(group_by(wtd01,discussing),
                             mean(learning.understanding)+
                             mean(success.understanding))
lu.split.read <- summarise(group_by(wtd01,reading..viewing),
                             mean(learning.understanding)+
                             mean(success.understanding))
lu.split.gpwk <- summarise(group_by(wtd01,groupwork),
                             mean(learning.understanding)+
                             mean(success.understanding))
lu.split.inwk <- summarise(group_by(wtd01,independent.work),
                             mean(learning.understanding)+
                             mean(success.understanding))
lu.split.livs <- data.frame(Category = "LI visible", 
                            Observation = c("No","Yes"), 
                            Rate = lu.split.livs[,2]/4)
lu.split.scvs <- data.frame(Category = "SC visible", 
                            Observation = c("No","Yes"), 
                            Rate = lu.split.scvs[,2]/4)
lu.split.drct <- data.frame(Category = "Direct Instruction", 
                            Observation = c("No","Yes"), 
                            Rate = lu.split.drct[,2]/4)
lu.split.disc <- data.frame(Category = "Discussion", 
                            Observation = c("No","Yes"), 
                            Rate = lu.split.disc[,2]/4)
lu.split.read <- data.frame(Category = "Reading / Viewing", 
                            Observation = c("No","Yes"), 
                            Rate = lu.split.read[,2]/4)
lu.split.gpwk <- data.frame(Category = "Group work", 
                            Observation = c("No","Yes"), 
                            Rate = lu.split.gpwk[,2]/4)
lu.split.inwk <- data.frame(Category = "Independent work", 
                            Observation = c("No","Yes"), 
                            Rate = lu.split.inwk[,2]/4)

mg <- merge(lu.split.livs,lu.split.scvs,all = TRUE)
mg <- merge(mg, lu.split.drct, all=TRUE)
mg <- merge(mg, lu.split.disc, all = TRUE)
mg <- merge(mg, lu.split.read, all = TRUE)
mg <- merge(mg, lu.split.gpwk, all = TRUE)
lu.split.cat <- merge(mg,lu.split.inwk, all = TRUE)
names(lu.split.cat) <- c("Category", "Observation", "Rate")

# Generating Plot
lu.split.plot <- nPlot(Rate ~ Category, group = "Observation",
                       type = "multiBarChart",
                       data = lu.split.cat)
#lu.split.plot$templates$script <- "http://timelyportfolio.github.io/rCharts_nvd3_templates/chartWithTitle.html"
lu.split.plot$set(title = "Understanding of learning process based on observations")
lu.split.plot$yAxis(tickFormat = "#!d3.format('%')!#")
lu.split.plot$chart(forceY = c(0,1))
lu.split.plot$yAxis(axisLabel = "Combined learning and success understanding")
lu.split.plot$print('iframe4',include_assets = TRUE)
```
