---
title: "Walkthough Analysis - Nitty Gritty"
author: "Ben Hicks"
output: html_document
---

Date: `r Sys.Date()`

This document is an informal collection of some of the details of the walkthroughs to date.


```{r echo=FALSE} 
# setting up library's and parameters
suppressPackageStartupMessages(library(googleVis))
suppressPackageStartupMessages(library(plyr))
op <- options(gvis.plot.tag='chart')

# reading in data, renaming columns
d0 <- data.frame(read.csv("Walkthrough.2016.02.10.csv"), 
                 Date = "2016-02-10")
d1 <- data.frame(read.csv("Walkthrough.2016.06.08.csv"),
                 Date = "2016-06-08")
d2 <- data.frame(read.csv("Walkthrough.2016.07.28.csv"),
                 Date = "2016-07-28")

# merging data
d <- rbind(d0,d1)
d <- rbind(d,d2)
wtd <- d
names(wtd) <- c('period','room','year.level','number.of.students','lesson.time','direct.instruction','discussing','reading..viewing','groupwork','independent.work','LI.visible','SC.visible','learning.understanding','success.understanding','date')

timeSeries <- ddply(wtd, ~date, summarise,
                    LI = mean(LI.visible),
                    SC = mean(SC.visible),
                    Learning.Understanding = mean(learning.understanding),
                    Success.Understanding = mean(success.understanding))
timeSeries$date <- as.Date(timeSeries$date, "%Y-%m-%d")
```

* `r length(wtd$period)` lessons observed so far in 2016.
* LI's visible `r round(100*mean(wtd$LI.visible))`% of the time and SC's visible `r round(100*mean(wtd$SC.visible))`% of the time
* Learning Understand averaged a score of `r round(50*mean(wtd$learning.understanding))`% and Success Understanding a score of `r round(50*mean(wtd$success.understanding))`%.

### Summary of activity:

``` {r echo = FALSE}
library(ggplot2)
act.dat <- data.frame(
  Activity = c("Direct Instruction", "Discussion", "Reading / Viewing", "Group Work", "Independent Work"),
  Frequency = c(sum(wtd$direct.instruction),
                sum(wtd$discussing),
                sum(wtd$reading..viewing), 
                sum(wtd$groupwork), 
                sum(wtd$independent.work))
)
ggplot(act.dat, aes(x = Activity, y = Frequency)) + geom_bar(stat = "Identity")

```

### Impact of LI, SC by lear level.

``` {r echo = FALSE}
suppressPackageStartupMessages(library(randomForest))

wtd.adj <- dplyr::mutate(wtd, Understanding = wtd$learning.understanding + wtd$success.understanding)
wtd.adj <- subset(wtd.adj, select = c(period,room,year.level,number.of.students,lesson.time,direct.instruction,discussing, reading..viewing, groupwork, independent.work, LI.visible, SC.visible, date, Understanding))


stage6 <- wtd.adj[wtd.adj$year.level %in% c(11,12),]
stage5 <- wtd.adj[wtd.adj$year.level %in% c(9,10),]
stage4 <- wtd.adj[wtd.adj$year.level %in% c(7,8),]
stage3 <- wtd.adj[wtd.adj$year.level %in% c(5,6),]
stage2 <- wtd.adj[wtd.adj$year.level %in% c(3,4),]
stage1 <- wtd.adj[wtd.adj$year.level %in% c(0,1, "K"),]

# Random forest method
set.seed(1234)
influence_all <- randomForest(Understanding ~ ., data = wtd.adj)
varImpPlot(influence_all)
influence_stage4 <- randomForest(Understanding ~ ., data = stage4)
varImpPlot(influence_stage4)
influence_stage5 <- randomForest(Understanding ~ ., data = stage5)
varImpPlot(influence_stage5)
influence_stage6 <- randomForest(Understanding ~ ., data = stage6)
varImpPlot(influence_stage6)

```


